Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-15 paper/main.tex
update

In many practical optimization problems, the objective function is not
explicitly available and solutions are evaluated by means of expensive
prediction models, simulations or physical experiments. When decision variables
are continuous, the use of Bayesian surrogate-models, e.g., Gaussian processes,
in optimization has become
widespread~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
success, there have been attempts at adapting such Bayesian optimization
algorithms to the combinatorial case, a notable example being Combinatorial
Efficient Global Optimization
(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
ruggedness of combinatorial landscapes, which makes local search particularly
effective in the white-box context, lessens the effectiveness of global
surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
expensive to train and optimize. Unless each function evaluation requires a day
or more, the time required by the Bayesian optimizer itself may impose a
significant overhead in computation time. \citet{PerLopStu2015si} recently
showed that ant colony optimization (ACO)~\citep{DorStu2004:book} is
competitive with CEGO on a black-box version of the travelling salesman problem
under a budget of $1\,000$ function evaluations. ACO basically builds a
probability distribution model from which solutions are sampled with a bias
towards the best solutions evaluated so far. As such, ACO may be considered a
type of estimation-of-distribution algorithm (EDA) for combinatorial
optimization.

@@

Te dejo aqui un comentario argumentando por qué he elegido estas instancias, puedes ver todas las graficas en el github

@@

\begin{tabular}{llrrr}
\toprule
Problem &                instance &  Fitness CEGO &  Fitness UMM &  Improvement \\
\midrule
    LOP &  lop/RandA1/N-t1d100.01 &       78696.6 &      76119.6 &     0.032746 \\
    LOP &  lop/RandA1/N-t1d100.02 &       79686.5 &      76827.7 &     0.035876 \\
    LOP &  lop/RandA1/N-t1d150.01 &      187605.5 &     179508.3 &     0.043161 \\
    LOP &  lop/RandA1/N-t1d150.02 &      184320.9 &     177592.8 &     0.036502 \\
    LOP &  lop/RandA1/N-t1d200.01 &      335391.1 &     323513.6 &     0.035414 \\
    LOP &  lop/RandA1/N-t1d200.02 &      333149.5 &     319675.8 &     0.040443 \\
    LOP &  lop/RandA2/N-t2d150.01 &       24682.7 &      15297.7 &     0.380226 \\
    LOP &  lop/RandA2/N-t2d150.02 &       24250.8 &      14724.9 &     0.392808 \\
   PFSP &          pfsp/rec05.txt &        1317.9 &       1330.7 &    -0.009712 \\
   PFSP &          pfsp/rec13.txt &        2128.2 &       2158.7 &    -0.014331 \\
   PFSP &          pfsp/rec19.txt &        2398.6 &       2405.1 &    -0.002710 \\
    QAP &           qap/kra32.dat &      115907.0 &     116320.0 &    -0.003563 \\
    QAP &           qap/nug12.dat &         624.2 &        660.4 &    -0.057994 \\
    QAP &           qap/nug30.dat &        7486.8 &       7474.0 &     0.001710 \\
    QAP &           qap/tho30.dat &      192416.0 &     191482.0 &     0.004854 \\
\bottomrule
\end{tabular}

>>>>

In many practical optimization problems, the objective function is not
explicitly available and solutions are evaluated by means of expensive
prediction models, simulations or physical experiments. When decision variables
are continuous, the use of Bayesian surrogate-models, e.g., Gaussian processes,
in optimization has become
widespread~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
success, there have been attempts at adapting such Bayesian optimization
algorithms to the combinatorial case, a notable example being Combinatorial
Efficient Global Optimization
(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
ruggedness of combinatorial landscapes, which makes local search particularly
effective in the white-box context, lessens the effectiveness of global
surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
expensive to train and optimize. The additional time required by the Bayesian
optimizer may impose a significant overhead in computation time, specially if
computation time of each function evaluation is measured in ``few'' minutes or
hours rather than days and when ``expensive'' refers to resources or economical
cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
ant colony optimization (ACO)~\citep{DorStu2004:book} is competitive with CEGO
on a black-box version of the travelling salesman problem under a budget of
$1\,000$ function evaluations. ACO basically builds a probability distribution
model from which solutions are sampled with a bias towards the best solutions
evaluated so far. As such, ACO may be considered a type of
estimation-of-distribution algorithm (EDA) for combinatorial optimization. On
the other hand, $1\,000$ evaluations is a relatively large budget for CEGO and
ACO was not designed for such short budgets.

@@

@@

Table~\ref{tab:results} shows

\begin{table}
\caption{Results\label{tab:results}}
\begin{tabular}{llrrr}
\toprule
Problem &                instance &  Fitness CEGO &  Fitness UMM &  Improvement \\
\midrule
    LOP &  lop/RandA1/N-t1d100.01 &       78696.6 &      76119.6 &     0.032746 \\
    LOP &  lop/RandA1/N-t1d100.02 &       79686.5 &      76827.7 &     0.035876 \\
    LOP &  lop/RandA1/N-t1d150.01 &      187605.5 &     179508.3 &     0.043161 \\
    LOP &  lop/RandA1/N-t1d150.02 &      184320.9 &     177592.8 &     0.036502 \\
    LOP &  lop/RandA1/N-t1d200.01 &      335391.1 &     323513.6 &     0.035414 \\
    LOP &  lop/RandA1/N-t1d200.02 &      333149.5 &     319675.8 &     0.040443 \\
    LOP &  lop/RandA2/N-t2d150.01 &       24682.7 &      15297.7 &     0.380226 \\
    LOP &  lop/RandA2/N-t2d150.02 &       24250.8 &      14724.9 &     0.392808 \\
   PFSP &          pfsp/rec05.txt &        1317.9 &       1330.7 &    -0.009712 \\
   PFSP &          pfsp/rec13.txt &        2128.2 &       2158.7 &    -0.014331 \\
   PFSP &          pfsp/rec19.txt &        2398.6 &       2405.1 &    -0.002710 \\
    QAP &           qap/kra32.dat &      115907.0 &     116320.0 &    -0.003563 \\
    QAP &           qap/nug12.dat &         624.2 &        660.4 &    -0.057994 \\
    QAP &           qap/nug30.dat &        7486.8 &       7474.0 &     0.001710 \\
    QAP &           qap/tho30.dat &      192416.0 &     191482.0 &     0.004854 \\
\bottomrule
\end{tabular}
\end{table}
